Multiphase implementation on a parallel 2D/3D Lattice Boltzmann solver using GPGPUs

### 2017-08-14 Thesis/Draft.tex
Conclusion done

\subsection{2D vs 3D}
We conclude this chapter by comparing the performance of the 2D model against the 3D model. A new metric must be defined to accurately show this, since it is obvious that under the same conditions the 3D model would take longer to process. We will show the results using MLUPS instead of seconds. MLUPS stands for Millions of nodes Updated Per Second.

@@

>>>>

\subsection{2D vs 3D}
We conclude this chapter by comparing the performance of the 2D model against the 3D model. A new metric must be defined to accurately show this, since it is obvious that under the same conditions the 3D model would take longer to process. We will show the results using MLUPS instead of seconds. MLUPS stands for Millions of nodes Updated Per Second, the higher the value, the better the performance.

    \begin{figure}[H]
    	\centering
    	\includegraphics[width=\linewidth]{Resources/Images/2dv3d.png}
    	\caption{MLUPS comparison between 2D and 3D model for 10000 iterations}
    	\label{fig:2dv3d}
    \end{figure}

These runs were executed using the solver's final version, compiled with each machine's appropriate CUDA compute capability, so as to show the solver's best performance in each GPU. We can clearly see that the 3D model benefits more from the solver's parallelization than the 2D model. In fact, we see an increase in performance of almost 5 times when comparing both model's best performance case. Unsurprisingly, this comes from the Grid in single precision, since this machine showed the best potential with our solver and single precision is usually faster than double precision on the GPU. 

@@

This project aimed to introduce multiphase capability into the previous in-house Lattice Boltzmann solver \cite{maciej_thesis,koleszar_thesis,jozsa_thesis,szoke_thesis}, specifically by using the Color Gradient model, both for 2D and 3D. \par
First, a literature review was performed, where the main focus areas surrounding the topic were explored. This was crucial to gain some insight on the theory of the Lattice Boltzmann. To this extent, we started by exploring the physical and mathematical side of the project, as well as understanding where the project could be applied to. Then we studied the technical side of the thesis, by exploring the history on parallel computing and subsequently of GPUs and General Purpose implementations. After this came the study on how the previous authors of the solver developed their work and what each of them purposed to implement. After gaining insight on all these areas, we were ready to start understanding the specifics of the methods. \par
We then presented our study on how each part of the Lattice Boltzmann method works, from the physical world to the actual software implementation. Our ideas regarding the actual implementation were already starting to form and as we progressed the implementation started becoming clearer. We also needed to gain some insight into how the Color Gradient model would function and how we could couple this multiphase implementation into the Lattice Boltzmann method. At this stage we could already perceive the types of meshes that we would need to use and how each of them would affect the final solution. Finally, before being able to parallelize the work, the CUDA platform was explored, ranging from it's underlying architecture to the available Macros to be used. \par
Having completed an in-depth study of the models, we were then ready to start with the implementation of the multiphase capabilities. However, before doing so, the previous solver needed to be analyzed in order to gain some knowledge on how to integrate the new features with the in-house code. We first started by implementing a serial version of the solver and only after producing correct, validated results were we comfortable with beginning the model's parallelization. Since we were implementing this in 2D and 3D, we had to repeat the process and start from the beginning to correctly implement the 3D version. Then it was a matter of extending the tests cases to cover the model's potential, as well as further validate the accuracy of our solver. After being satisfied with the results obtained so far, we shifted our attention from the physical results to the performance results. We started by analyzing the solver's memory usage to try and extend the size of the problems being solved, but could not find a way to do this without sacrificing the solver's accuracy. Then we performed an extensive study on the model's behavior in terms of run-times. Each of project's main improvements was documented and tested for both precisions and on all available machines. The final version provided some good results, where we could see improvements of up to almost 18\% in 2D and 3D when comparing with the initial parallel version. This parallelization translated into an increased performance of up to 20 times for 2D and 176 times for 3D when compared with the initial serial version. We could also a better potential for the model in 3D problems than on 2D problems since the performance was 5 times better for the former.\par
However, there is much work to be done. Time was a big constraint on this project since introducing new features into an already existing solver proved to be quite a daunting task. There are many unexplored capabilities for the multiphase implementation, like for example introducing new multiphase models or even changing the collision model in the CG implementation to a multi-relaxation time approach. There is even room for further improving the software side of the solver, like making the 2D model more flexible in terms of mesh interpolation or further increasing the performance of the system by employing different computation or memory storage techniques, like the usage of CUDA's texture memory for example. One particular important aspect would be to implement better memory usage so as to increase the size of problems to be computed. \par
Nonetheless, this was a very exciting and challenging project for our thesis since it provided the opportunity of merging software development with engineering problems, while using newer and different techniques, namely parallel programming under the CUDA platform. There was also a big satisfaction in overcoming the initial barrier of the physical and mathematical demands of this type of solvers and finally being able to produce correct and visual solutions, making it a very interesting topic for a Master's thesis.
